---
title: "regression"
author: "Kyle Bistrain"
format:
  html:
    self-contained: true
    mainfont: Times New Roman
editor: visual
execute:
 echo: true
 error: true
 warning: true
---

```{r}
#| warning: false
library(tidyverse)
library(broom)

```

1\.

```{r}
#| warning: false
save <- read.csv(here::here('supporting_artifacts', 
                           'learning targets', 
                           'mystery_animal.csv'
                           )
                 )
```

2\.

```{r}

save |> 
  lm(weight_after ~ weight_before, data = _) |> 
  augment() |> 
  ggplot(aes(x = weight_before, y = .resid)) +
  geom_point()


```

3\.

```{r}

trumpet <- runif(100,1.5,3.5)
qunif(p = .95, min = 1.5, max = 3.5 )
```

-   4.What is the 95th percentile for trumpet weight?

-   5.What is the 10th percentile for trombone weight?

    ```{r}
    qnorm(p =.1, mean =  4.6,sd= .8)

    ```

-   6.About what percent of trombones do you expect to be more than 5 pounds?

    ```{r}
    1- pnorm(5, mean = 4.6, sd = .8)
    ```

-   7.About what percent of reed instruments do you expect to be more than 5 pounds?

    ```{r}
    1-pchisq(5, df = 4 )
    ```

-   8.Simulate 100 random trombone weights. How many of them were below 4 pounds?

```{r}
a <- rnorm(n =100, mean = 4.6, sd = .8)

sum (a < 4 )
```

# **3 Catching a Con**

You manage to intercept a shipping notice for delivery to Professor Hill. It says the following:

    Wells Fargo Shipment 1957

    To:  Harold Hill, Mason City, Iowa
    From:  Music Factory, Gary, Indiana

    Items included:
        Trombones: 76
        Cornets: 110
        Saxophones: 542
        Clarinets: 318
        Bassoons: 175
        
    Total Shipped Weight: 4532 lbs.

This is your chance to catch Professor Hill in his lie!

First, write a function that samples the correct number of weights of trombones, cornets (trumpets), and reed instruments (saxophones, clarinets, bassoons), and then returns the total weight of the shipment.

Then, use that function to create random samples of the total weight of 1000 possible shipments.

(Some code is provided for you below. Do not change the seed!)

How many of these samples had a weight less than Prof. Hill's shipment? Do you believe Professor Hill is lying, or not?

```{r}


music_man <- function(n_tromb, n_cor, n_reed){
  
  trombones <- rnorm(n = n_tromb, mean = 4.6, sd = .8)
  cornets <- runif(n = n_cor, min = 1.5, max = 3.5)
  reeds <- rchisq(n = n_reed, df = 4)
  
  return(sum(trombones,cornets,reeds))
  
}

set.seed(1957)

my_weights <- rerun(.n = 1000, 
                    music_man(n_tromb = 76,n_cor =  110, n_reed = 1035)
                    ) |> 
  flatten_dbl()

```

```{r}
sum(my_weights<4532)
```

He is lying because only 7 shipments are less than 4532 pounds. therefore there is a low probability of this occuring. 7/1000 = .007. aproximate p-value of .007
